### 2020-04-25 generated/paper_EN.Rmd
Add generated files

---
title:
  Concrete compressive strength prediction with machine learning
date: "April 22, 2020"
author: "Pedro Bernardino Alves Moreira"
indent: true
abstract:
 Compressive strength is the main characteristic of concrete. This work used a set of samples widely used to build 6 forecast models of this characteristic, each one specifically for an age, 3, 7, 14, 28, 56 and 100 days from 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers) and 6 relationships between the main ingredients (water/cement, coarse aggregate/cement, fine aggregate/cement, fine aggregate / coarse aggregate, water / coarse aggregate and water/fine aggregate) to make the predictions. The algorithm *Parallel Random Forest* was used (*parRF* of the *caret* package in *R*) . The results were satisfactory, being the best for the 3-day model (*RMSE* of 3.31) and the worst for the 100-day model (*RMSE* of 5.85). The 28-day model had *RMSE* of 4.72. The work indicated that the separation in a model for each age is promising and was able to generate compatible results or better than other studies, opening way for new models creation following this proposal, but using different algorithms or the combination of different algorithms.
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
bibliography: references.bib
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{supertabular}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{caption}
- \usepackage{floatrow}
- \floatsetup[table]{capposition=top}
- \floatsetup[figure]{capposition=top}
- \captionsetup{options=chunk}
- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
- \renewcommand{\thechunk}{\arabic{chunk}}
- \usepackage{indentfirst}
- \usepackage{sectsty} \sectionfont{\centering}
---
```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")
# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)
# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
```
\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

# Introduction
Compressive strength is the main characteristic of concrete, measured by tests of international standards that consist of the breaking of specimens. Measurement at 28 days is mandatory and represents the grade of the concrete. Knowing in advance what the result will be obtained for a given age, based on the proportions of its ingredients, is of great interest to concrete manufacturers, construction companies, and civil engineers.

&nbsp;

This compressive strength is a nonlinear function of its ingredients and age, making it difficult to establish an analytical formula, although some formulas have already been proposed. @Hasan2011 proposed a mathematical model to predict from the results of tests of 7 and 14 days, and @Kabir2012 from 7 days. However, machine learning techniques can be used to model this characteristic from real sample data, using only the ingredients.

&nbsp;

Many previous studies use the same dataset used by @Yeh1998 to predict the compressive strength of concrete. @Alshamiri2020 got good results with the regularized extreme learning machine (RELM) technique, and @Hameed2020 got even better results with the Artificial Neural Networks and cross-validation technique. This set of samples is so well known that there are many pages on the internet of unpublished studies that use it and have good results, such as @Abban2016, @Raj2018, @Modukuru2020 and @Pierobon2018. At the end of the work, the results found are compared to the works cited here.

&nbsp;

Unlike previous studies with this dataset, this work does data preparation differently. The age of the concrete is the most unique feature that contributes to its compressive strength. For this reason, age is treated separately in the machine learning models, creating models for each age group.

# Materials and Methods

## Materials
The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore]. Throughout the process, all code executed was documented in the same order as its execution in [Appendix 3](#appendix3), and reference was always made to codes throughout the text. All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1). In addition, an online and open-source repository was created in *Github*, housing all the code used to generate this work, the link was made available in [Appendix 2](#appendix2).

## Reproducibility
In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result. The *seeds* can be checked throughout [Appendix 3](#appendix3) or directly on *Github*.

## Obtaining the data
The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}). In total there are 1030 samples with 9 columns. The samples were renamed and an id column was added to facilitate data manipulation (\ref{show-rename-dat-cols}). The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}). The first samples can be viewed in the table \ref{tab: first-samples}.

&nbsp;

```{r download-data, echo=F, include=F}
# Data download
url_base <- "https://archive.ics.uci.edu"
url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```
```{r rename-dat-cols, echo=F}
# Ranaming the columns
colnames(dat) <- c(
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
  )
dat$id <- seq.int(nrow(dat))
```
```{r reorder-dat, echo=F}
# Reordering the data
col_order <- c(
  "id",
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
)
dat <- dat[, col_order]
```
```{r col-names-and-units, echo=F}
# Defining column names and units
colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
```
```{r first-samples, echo=F}
# Table First samples
caption <- "First 6 samples"
kable(
    dat[1:6,],
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

## Data preparation
The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies. Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed. In addition, the relevant data has been transformed to better fit the studies in the next steps.

### Initial data cleaning
Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).

&nbsp;

```{r removing-duplicated-samples, echo=F}
# Removing duplicated samples
n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
n_duplicated_samples <- n_inicial_samples - n_distinct_samples
dat <- dat[!duplicated(select(dat, -c(id))),]
n_samples <- nrow(dat)
```
The data show the variables in the columns and samples in the rows. However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strngth, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.

&nbsp;

```{r similar-samples, echo=FALSE}
# Tabela - Samples with same composition
same_samples <- dat %>%
  filter(id %in% c(653, 678, 654, 681))
caption <- "Samples with same composition"
kable(
  same_samples[order(same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process. This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.

&nbsp;

```{r similar-samples-2, echo=FALSE}
# Tabela - Same samples with different results
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Same samples with different results"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```
To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients (\ref{show-total-de-amostras-apos-processamento}).

&nbsp;

```{r initial-data-cleaning, echo=FALSE}
# Data cleaning
dat <- dat %>%
  group_by(
    cement,
    blast_furnace_slag,
    fly_ash,
    water,
    superplasticizers,
    coarse_aggregate,
    fine_aggregate,
  ) %>%
  filter("28" %in% day) %>%
  mutate(id = id[which.min(id)]) %>%
  ungroup() %>%
  group_by(id, day) %>%
  mutate(mpa = mean(mpa)) %>%
  ungroup()
dat <- dat[!duplicated(select(dat, -c(id))),]
dat$id<-factor(dat$id)
n_samples <- nrow(dat)
n_distinct_samples <- n_distinct(dat$id)
```
The result can be seen in the table \ref{tab:similar-samples-same-id}. All samples with equal ingredient settings have the same id, and when they had different results for the same days, they were transformed into just one sample, with the arithmetic mean in the compressive strength.

&nbsp;

```{r similar-samples-same-id, echo=FALSE, include=TRUE}
# Tabela - Previous samples after processing
same_samples <- dat %>%
  filter(id == 653 | id == 472 & day == 28)
caption <- "Previous samples after processing"
kable(
  same_samples[order(same_samples$id, same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE,
  digits = 2
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

### Age selection
As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build forecasting models. However, it is necessary to verify how relevant the data of these other ages are. Starting with the distribution of the samples in relation to each age  (\ref{show-boxplot}) shown in the figure \ref{fig:boxplot}.

&nbsp;

```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
# Figure - Compressive strength (MPa) vs age (days)
cap <- "Boxplot - Compressive strength (MPa) vs age (days)"
ylabel <- "Compressive strength (MPa)"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x=factor(day), y=mpa)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)  +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel)
```
It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.

&nbsp;

This hypothesis was verified using the principal component analysis method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}). The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis. The first two dimensions represent $ 37 \% $ and $ 24 \% $ respectively of the variance.

&nbsp;

```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
# Figure - Principal component analysis - 90, 91 e 100 days
dat_90_91_100 <- dat %>%
  ungroup() %>%
  filter(day %in% c(90, 91, 100)) %>%
  select(-c(id, mpa))
cap <- "Principal component analysis - 90, 91 e 100 days"
colnames(dat_90_91_100) <- c(
   "Cem.", "B.F.S.", "Fly.A.","Water","Sup.","C.Ag.","F.Ag.","day")
pca <- prcomp(select(dat_90_91_100, -c(day)), scale = TRUE)
habillage <- dat_90_91_100$day
fviz_pca_biplot(
  pca,
  geom.ind = "point",
  habillage=habillage,
  addEllipses = TRUE,
  ellipse.level=0.75) +
  ggtitle("") +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
```
Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value. The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-team}).

&nbsp;

```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Compressive strength through time
cap <- "Compressive strength through time"
ylabel <- "MPa"
xlabel <- "Age (days)"
dat_duplicated_only <- dat %>%
  group_by(id) %>%
  filter(n()>5) %>%
  select(id, mpa, day)
dat_duplicated_only %>%
  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
  geom_line() +
  xlab(xlabel) +
  ylab(ylabel) +
  theme_bw() +
  theme(legend.position = "none")
```
For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise. They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}. As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.

&nbsp;

```{r join-90-91-100, echo=F, eval=F}
# Joining 90, 91 and 100 days data
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```
Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.

&nbsp;

```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Ages frequency
cap <- "Ages frequency"
ylabel <- "Frequency"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x = factor(day))) +
  geom_bar() +
  theme_bw() +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r remove-ages-lower-50, echo=F}
# Removing ages with frequency lower than 50
dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
```

### Data reorganization
The samples were grouped to maintain only a separate sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}). The result in the first samples after this processing is shown in the table \ref{tab:new-features}.

&nbsp;

```{r reorganizing-dat, echo=F}
# Data reorganization
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```
```{r new-features, echo=F}
# Table - First 6 samples after reorganization/EN>
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).
```{r total-samples-2, echo=F}
# Total samples/EN>
n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

### Adding new variables
To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table ref{tab:new-features-table}.
```{r new-features-2, echo=F}
# Adding New features/EN>
concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}
mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}
dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```
```{r new-features-table, echo=F}
# Table - New features/EN>
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

## Data visualization
In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.

### Descriptive statistics
The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ ref {show-stat-summ}). The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data. As the samples were filtered to maintain only sets of samples with known values of compressive strength at 28 days, the number of *NAs* is zero for that age. The figure \ref{fig:stat-summ-categorical} presents the statistical data of the discrete variables (\ref{show-stat-summ-categorical}).

&nbsp;

```{r stat-summ, echo=FALSE}
# Table - Descriptive statistics - continuous variables/EN>
summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```
```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
# Figure - Descriptive statistics - categorical variables/EN>
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

### Correlation between ingredients and compressive strength
The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}). The figure \ref{fig:correlation-mpa} presents the same data, but instead of correlating them all, it only correlates with the compressive strength, showing the values in more detail (\ref{show-correlation-mpa}).

&nbsp;

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
# Figure - Correlation grouped by age/EN>
cor_dat <- dat %>% select(-c(id))
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.", "3", "7",
                "14", "28", "56", "100",
                "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```
The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.

&nbsp;

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
# Figure - Relationship between approximated mix, water, MPa and age/EN>
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
mix_dat$day <- factor(mix_dat$day)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
# Figure - Relationship between concrete main features/EN>
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```
The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

### Variables distribution
A figura \ref{fig:vars-distribution} mostra a distribuição das variáveis nas amostras (\ref{show-vars-distribution}). Foi calculado utilizando apenas os dados aos 28 dias. 
The figure\ref{fig:vars-distribution} shows the distribution of variables in the samples (\ref{show-vars-distribution}). It was calculated using data only at 28 days.

&nbsp;

```{r vars-distribution, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message = F}
# Figure - Variables distribution/EN>

cap <- "Variables distribution"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
dist_dat <- dat %>%
  select(-c(id, class, day_3, day_7, day_14, day_56, day_100, mix_app))

colnames(dist_dat) <- colnames_dat
dist_dat <- dist_dat %>%
  gather("Var", "value") %>%
  mutate(value = as.numeric(value))
dist_dat %>%
  ggplot(aes(value)) +
  geom_histogram(aes(y = ..density..), 
                 colour = "black", 
                 fill = "white") +
  geom_density(alpha = .5, fill = "lightseagreen") +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  xlab("") +
  ylab("") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```
The figure \ref{fig:vars-distribution-time} shows the distribution of ingredients and compressive strength for each set of ages (\ref{show-vars-distribution-time}), in case of the 28 days it presents the same information as the figure \ref{fig:vars-distribution}. It shows that the resistance to compression gradually increases over time, as expected. Furthermore it is seen that the concentration of the ingredients can vary a lot when stratified by ages.
```{r vars-distribution-time, echo=F, fig.height=10, fig.cap=cap, warning=FALSE, message=F}
# Figure - Variables distribution grouped by age/EN>

days_labs <- c("3","7","14","28","56","100")
cap <- "Variables distribution grouped by age"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                  "Superplast.", "Coarse agg.", "Fine agg.",
                  days_labs,
                  "Water / Cement", "Fine agg. / Cement", 
                  "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
                  "Water / Coarse agg.", "Water / Fine agg.")
dist_dat_2 <- dat %>%
  select(-c(id, class, mix_app))
colnames(dist_dat_2) <- colnames_dat
dist_dat_2 <- dist_dat_2 %>%
  gather("day", "mpa", days_labs) %>%
  drop_na() %>%
  gather("Var", "value", -c("day")) %>%
  mutate(day = factor(day, levels = days_labs))
dist_dat_2 %>%
  ggplot(aes(x = day, y = value)) +
  geom_violin(color = NA,
                fill = "lightseagreen",
                alpha = .5,
                na.rm = TRUE,
                scale = "count") +
  geom_boxplot(alpha = 0.2) +
  facet_wrap(~ Var, ncol=3, scale = "free") +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black")) +
  xlab("") +
  ylab("")
```

### Principal component analysis
Na figura \ref{fig:pca}, utilizando uma classificação alternativa, foi realizada a análise de componente principal nos ingredientes das amostras (\ref{show-pca}). A classificação separa o concreto em 4 grupos de resistência a compressão diferentes, baixo até $20 MPa$, normal até $40 MPa$, médio até $70 MPa$ e alto acima disso. É possivel perceber que os grupos se sobrepõem, mas existem uma diferenciação entre o grupo alto e baixo.
In the figure v, using an alternative classification, the principal component analysis was performed on the ingredients (\ref{show-pca}). The classification separates concrete into 4 different compressive strength groups, low up to *20 MPa*, normal up to *40 MPa*, medium up to *70 MPa* and high above that. It is possible to notice that the groups overlap, but there is a differentiation between the high and low group.
```{r pca, echo=F, fig.height=3, fig.cap=cap}
# Figure - Principal component analysis on ingredients/EN>

cap <- "Principal component analysis on ingredients"
colnames_dat <- c(
  "Cem.", "B.F.S.", "F.A.", "Wat.", "Sup.", "C.Agg.",
  "F.Agg.","WxC", "F.Agg.xC","C.Agg.xC","F.Agg.xC",
  "WxC.Agg.","WxF.Agg."
)
class_list <- list(low="0", normal="20", medium="40", high="70")
dat_pca <- dat %>%
  select(-c(id, class, mix_app,
            "day_3", "day_7", "day_14", "day_28", "day_56", "day_100"))
colnames(dat_pca) <- colnames_dat
class_2 <- dat$day_28
class_2[class_2 < 20] = "0"
class_2[class_2 >= 20 & class_2 < 40] = "20"
class_2[class_2 >= 40 & class_2 < 70] = "40"
class_2[class_2 >= 70] = "70"
class_2 <- factor(class_2)
levels(class_2) <- class_list
pca <- prcomp(dat_pca, scale = TRUE)
fviz_pca_ind(
  pca,
  geom.ind = "point",
  habillage=class_2,
  addEllipses = T,
  ellipse.level=0.95) +
  ggtitle("") +
  theme_bw()
```

## Machine learning models
The development of machine learning models was carried out with the *caret* package [@caret] and based on @irizarry2019 e @Kuhn2008.

### Pre-processing and data separation
Como existe a variável categórica para o traço aproximado do concreto, foi realizada a conversão dessa variável em variáveis fictiícias (*dummy vars*) (\ref{show-dummy-var}), passando de 22 colunas (id, classe, resistência à compressão e mais 19 *features*) para 45 colunas, uma adição de 23 variáveis, uma para cada traço aproximado. 
As the approximate mix is a categorical variable, it was converted into dummy variables (\ref{show-dummy-var}), going from 22 columns (id, class, compressive strength and 19 more *features*) to 45 columns, an addition of 23 variables, one for each approximate mix.

&nbsp;

```{r dummy-var, echo=F, message=F, warning=F}
# Dummy variables
dummies <- dummyVars( ~ mix_app, data = dat)
dummyDat <- data.frame(predict(dummies, newdata = dat))
dummyDat$id <- dat$id
dat <- dat %>%
  select(-c(mix_app)) %>%
  full_join(., dummyDat)
```
The samples were separated based on age. One dataset was created for each age value, totaling 6 different sets (\ref{show-preparation}). For illustrative purposes, the first 18 of 45 columns of the first 6 samples from the 28-day set are shown in the table \ref{tab:table-prepared-samples}.

&nbsp;

```{r preparation, echo=F}
# Data preparation
names(dat) <- gsub(x = names(dat), pattern = "/", replacement = ".")  
dat_3 <- dat %>%
  select(-c("day_7", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_3",~"mpa")
dat_7 <- dat %>%
  select(-c("day_3", "day_14", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_7",~"mpa")
dat_14 <- dat %>%
  select(-c("day_3", "day_7", "day_28", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_14",~"mpa")
dat_28 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_56", "day_100")) %>%
  drop_na() %>%
  rename_at("day_28",~"mpa")
dat_56 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_100")) %>%
  drop_na() %>%
  rename_at("day_56",~"mpa")
dat_100 <- dat %>%
  select(-c("day_3", "day_7", "day_14", "day_28", "day_56")) %>%
  drop_na() %>%
  rename_at("day_100",~"mpa")
```
```{r table-preparated-samples, echo=F}
# Table - First 18 columns of the 6 first samples of 28 days
colNames = c("ID", "Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.", "MPa", "Class",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "App Mix" = 2)
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","", "Ci.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "1:1:2", "1:2:2")
caption <- "First 18 columns of the 6 first samples of 28 days"
dat_tabela <- dat_28[1:18]
kable(
    head(dat_tabela[order(dat_tabela$id),]),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
For each of the 6 sets, the existence or not of variables with variance close to zero and their subsequent removal was verified (\ref{show-nzv}). Many of the 23 variables added referring to the approximate concrete mix were removed due to this fact. In addition to them, in the case of the 7-day set, the fly ash variable was also removed. Then it was verified that there are no variables with high correlation, above 0.999, in any of the 6 data sets (\ref{show-cors}). After these steps, the sample sets presented 24, 21, 23, 23, 24 and 25 columns respectively for ages in increasing sequence.

&nbsp;

```{r nzv, echo=F}
# Removing near zero variance columns
nzv_3 <- nearZeroVar(dat_3)
nzv_7 <- nearZeroVar(dat_7)
nzv_14 <- nearZeroVar(dat_14)
nzv_28 <- nearZeroVar(dat_28)
nzv_56 <- nearZeroVar(dat_56)
nzv_100 <- nearZeroVar(dat_100)
dat_3 <- dat_3[,-nzv_3]
dat_7 <- dat_7[,-nzv_7]
dat_14 <- dat_14[,-nzv_14]
dat_28 <- dat_28[,-nzv_28]
dat_56 <- dat_56[,-nzv_56]
dat_100 <- dat_100[,-nzv_100]
```
```{r cors, echo=F}
# High correlation variables verification
descr_cor_3 <- cor(select(dat_3, -c(mpa, id, class)))
descr_cor_7 <- cor(select(dat_7, -c(mpa, id, class)))
descr_cor_14 <- cor(select(dat_14, -c(mpa, id, class)))
descr_cor_28 <- cor(select(dat_28, -c(mpa, id, class)))
descr_cor_56 <- cor(select(dat_56, -c(mpa, id, class)))
descr_cor_100 <- cor(select(dat_100, -c(mpa, id, class)))
high_cor_3 <- sum(abs(descr_cor_3[upper.tri(descr_cor_3)]) > .999)
high_cor_7 <- sum(abs(descr_cor_7[upper.tri(descr_cor_7)]) > .999)
high_cor_14 <- sum(abs(descr_cor_14[upper.tri(descr_cor_14)]) > .999)
high_cor_28 <- sum(abs(descr_cor_28[upper.tri(descr_cor_28)]) > .999)
high_cor_56 <- sum(abs(descr_cor_56[upper.tri(descr_cor_56)]) > .999)
high_cor_100 <- sum(abs(descr_cor_100[upper.tri(descr_cor_100)]) > .999)
```
The stage of centralization and normalization of the variables was carried out later, together with the application of the models, as it is simpler to do this with the *caret* package. If performed at this time, it would be necessary to manually undo these transformations in the predictions. The *caret* allows you to transform before training the models and already transforms the results back.

&nbsp;

Each data set was separated into test and training sets, *20%* and *80%* respectively (\ref{show-split}). The figure \ref{fig:dist-split} shows the distribution of data between the sets in relation to the compressive strength for each model (\ref{show-dist-split}).
```{r split, echo=F, message=F, warning=F}
# Test and train data split
reg <- list(
  dat_3 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_3$MPa),
  dat_7 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_7$MPa),
  dat_14 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_14$MPa),
  dat_28 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_28$MPa),
  dat_56 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_56$MPa),
  dat_100 %>% select(-c(mpa, class, id)) %>% mutate(y = dat_100$MPa)
)
reg_seed <- c(
  1111, # 3
  1, # 7
  22, # 14
  11111, # 28
  111, # 56
  11 # 100
  )
split_reg <- function(n){
  set.seed(reg_seed[[n]], sample.kind="Rounding")
  createDataPartition(reg[[n]]$y, p = .8, list = F)
}
trainIndex_reg <- lapply(list(1,2,3,4,5,6), split_reg)
gen_dat <- function(n){
  regIndex <- trainIndex_reg[[n]]
  list(train = reg[[n]][regIndex,], test = reg[[n]][-regIndex,])
}
dats_reg <- lapply(list(1,2,3,4,5,6), gen_dat)
names(dats_reg) <- c("d3", "d7", "d14", "d28", "d56", "d100")
```
```{r dist-split, echo=F, message=F, warning=F, fig.height=3, fig.cap= cap}
# Distribution of test and train data
cap <- "Distribution of test and train data"
d_lab <- " days"
ylabel <- "Density"
g1 <- "Train"
g2 <- "Test"
dens <- function(d, n){
  dens <- full_join(
      d$train %>%
        select(y) %>%
        mutate(Grup = g1, day = str_sub(n, 2)),
      d$test %>%
        select(y) %>%
        mutate(Grup = g2, day = str_sub(n, 2))
    )
  dens
}
densities <- imap(dats_reg, dens) %>%
  reduce(rbind) %>%
  mutate(day_f = factor(day, levels=c('3','7','14','28', '56', '100')))
labs <- paste(c("3","7","14","28","56","100"), d_lab, sep="")
levels(densities$day_f) <- labs
densities %>%
  ggplot(aes(y, fill = Grup)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~dia_f, ncol=3) +
  xlab("MPa") +
  ylab(ylabel) +
  theme_bw() +
  theme_minimal() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

### Performance measures
The performance evaluation of the models was performed by the Root Mean Square Error (*RMSE*). The *RMSE* is the measure used in all the works mentioned in the introduction and will allow the comparison of the models in the discussion.

### Naive models
Before creating the real models, for comparison purposes, naive models were created. They simply predict that the compressive strength of the test set is the average compressive strength of the training set (\ref{show-naive-model-reg}). In other words, naive models are simply the best guess possible. The results can be checked in the table \ref{tab:table-naive-model-reg}.
```{r naive-model-reg, echo=F, message=F, warning=F}
# Naive model
res <- function(d, n){
  data.frame(
    day = str_sub(n, 2),
    mean_mpa = mean(d$train$y)   
    )
}
ing_model <- imap(dats_reg, res)
get_rmse <- function(d, n){
   data.frame(
    rmse_train = RMSE(d$train$y, ing_model[[n]]$mean_mpa),
    rmse_test =  RMSE(d$test$y, ing_model[[n]]$mean_mpa),
    day = str_sub(n, 2) 
    )
}
rmses <- imap(dats_reg, get_rmse) %>%
  reduce(rbind)
ing_model_df <- ing_model %>% reduce(rbind)
df_performance_reg <- full_join(ing_model_df, rmses)
```
```{r table-naive-model-reg, echo=F}
# Table - Naive models
colNames = c("Age", "Mean $MPa$ (train)", "RMSE (train)", "RMSE (test)")
caption <- "Naive models"
kable(
    df_performance_reg,
    col.names = colNames6,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```

### Choice of algorithms
The *caret* [@caret] package exposes more than 200 different algorithms for building *machine learning* models. The package documentation presents an initial code [@modelsClusters] as a suggestion to select a portfolio of the most distinct algorithms possible in relation to some pre-selected algorithm, but for agility and due to technical limitations, it was chosen to use an algorithm with the highest probability to achieve the best possible result. According to @Fernandez2014, who compared 179 algorithms in 121 different databases, the algorithm most likely to achieve the best possible results is the *Parallel Random Forest* (called *prRF * in *caret*).

### Regression models
As new variables were added throughout the processing (the relationships between the ingredients and a few more *dummy vars* for each age set), 5 possibilities for configuring the *features* for the models were studied:
1. All *features*;
2. All *features* without the *dummy vars*;
3. Only the original *features*;
4. Only new *features*;
5. Only new *features*, without *dummy vars*;
Building a model for each of these configurations using the set at 28 days (\ref{show-test-models-reg}), showed that the best option is configuration 2, the *dummy vars* were completely discarded, but the other new variables were kept.
For illustrative purposes, the \ref{tab:table-done-samples-28} table shows the first 6 samples from the 28-day model training set. The samples of the other models, of the test and training sets are similar, the only difference being in the 7-day model, which excludes fly ash due to the variance close to zero, performed previously.
```{r test-models-reg, echo=F, warning=F, message=F, include=F}
# Features selection
data1 <- dats_reg$d28$train # full
data2 <- dats_reg$d28$train[c(1:13, 21)] # no dummy vars
data3 <- dats_reg$d28$train[c(1:7, 21)] # only original variables
data4 <- dats_reg$d28$train[c(8:21)] # only new variables
data5 <- dats_reg$d28$train[c(8:13, 21)] # only new varibels, no dummy vars
test1 <- dats_reg$d28$test # full
test2 <- dats_reg$d28$test[c(1:13, 21)] # no dummy vars
test3 <- dats_reg$d28$test[c(1:7, 21)] # only original variables
test4 <- dats_reg$d28$test[c(8:21)] # only new variables
test5 <- dats_reg$d28$test[c(8:13, 21)] # only new varibels, no dummy vars
# parRF
modelLookup("parRF")
control_parRF <- trainControl(method='repeatedcv',number=10,repeats=5,search='grid')
fit_parRF<- function(data, n) {
  set.seed(1, sample.kind = "Rounding")
  tunegrid_parRF<- expand.grid(mtry = seq(1, length(data), n))
  train(y ~ ., 
    data = data,
    preProcess = c("center","scale"),
    method='parRF', 
    tuneGrid=tunegrid_parRF,
    trControl=control_parRF)
}
fit_parRF1 <- fit_parRF(data1, 3)
ggplot(fit_parRF1)
min(fit_parRF1$results$RMSE) # 6.26108
p1_parRF1 <- predict(fit_parRF1, newdata = test1)
RMSE(p1_parRF1, test1$y) # 4.887264
fit_parRF2 <- fit_parRF(data2, 1)
ggplot(fit_parRF2)
min(fit_parRF2$results$RMSE) # 6.268789
p_parRF2 <- predict(fit_parRF2, newdata = test2)
RMSE(p_parRF2, test2$y) # 4.812994
fit_parRF3 <- fit_parRF(data3, 1)
ggplot(fit_parRF3)
min(fit_parRF3$results$RMSE) # 6.361905
p_parRF3 <- predict(fit_parRF3, newdata = test3)
RMSE(p_parRF3, test3$y) # 5.083467
fit_parRF4 <- fit_parRF(data4, 1)
ggplot(fit_parRF4)
min(fit_parRF4$results$RMSE) # 8.705984
p_parRF4 <- predict(fit_parRF4, newdata = test4)
RMSE(p_parRF4, test4$y) # 8.083962
fit_parRF5 <- fit_parRF(data5, 1)
ggplot(fit_parRF5)
min(fit_parRF5$results$RMSE) # 8.691755
p_parRF5 <- predict(fit_parRF5, newdata = test5)
RMSE(p_parRF5, test5$y) # 7.991537
```
```{r table-done-samples-28, echo=F}
# Table - First 6 samples of train data of the 28 days model
caption <- "First 6 samples of train data of the 28 days model"
colNames = c("Cement", "B.F.S.", "Fly ash", "water",
              "Superp.", "Coarse Agg.", "Fine Agg.",
              "Wat./", "F.Agg./", "C.Agg./",
              "F.Agg./","Wat./", "Wat./", "y")
dfUnits <- c("$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","Ce.",
              "Ce.", "Ce.", "C.Agg.", "C.Agg.", "F.Agg.", "MPa")
colNames2 = c("Features"=13, "Outcome"=1)
kable(
    head(data2),
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    digits = 2,
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    add_header_above(header = colNames2, line = T, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```
For each age set, a model was created using the *Parallel Random Forest* algorithm, previously defined (\ref{show-reg-models}). For each of the 6 models, the parameter *mtry* was optimized, and *repeated cross-validation * was performed, dividing into 10 or 30 parts and repeating 10 times.
```{r reg-models, echo=F, warning=F, message=F, include=F}
# Regression models
# no dummy vars:
data3 <- dats_reg$d3$train[c(1:13, 22)]
data7 <- dats_reg$d7$train[c(1:12, 19)]
data14 <- dats_reg$d14$train[c(1:13, 21)]
data28 <- dats_reg$d28$train[c(1:13, 21)]
data56 <- dats_reg$d56$train[c(1:13, 22)]
data100 <- dats_reg$d100$train[c(1:13, 23)]
test3 <- dats_reg$d3$test[c(1:13, 22)]
test7 <- dats_reg$d7$test[c(1:12, 19)]
test14 <- dats_reg$d14$test[c(1:13, 21)]
test28 <- dats_reg$d28$test[c(1:13, 21)]
test56 <- dats_reg$d56$test[c(1:13, 22)]
test100 <- dats_reg$d100$test[c(1:13, 23)]
get_trControl <- function(n, r){
  trainControl(method='repeatedcv',number=n,repeats=r,search='grid')
}
trControl3 <- get_trControl(30, 10)
trControl7 <- get_trControl(10, 10)
trControl14 <- get_trControl(30, 10)
trControl28 <- get_trControl(30, 10)
trControl56 <- get_trControl(30, 10)
trControl100 <- get_trControl(10, 10)
get_tuneGrid <- function(data){
  expand.grid(mtry = seq(1, length(data), 1))
}
tuneGrid3 <- get_tuneGrid(data3)
tuneGrid7 <- get_tuneGrid(data7)
tuneGrid14 <- get_tuneGrid(data14)
tuneGrid28 <- get_tuneGrid(data28)
tuneGrid56 <- get_tuneGrid(data56)
tuneGrid100 <- get_tuneGrid(data100)
set.seed(1, sample.kind = "Rounding")
fit_3 <- train(y ~ ., 
            data = data3,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid3,
            trControl=trControl3)
p_3 <- predict(fit_3, newdata = test3)
RMSE_test_3 <- RMSE(p_3, test3$y)
RMSE_test_3 # 3.31037
fit_3$bestTune # mtry = 6
set.seed(1, sample.kind = "Rounding")
fit_7 <- train(y ~ ., 
            data = data7,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid7,
            trControl=trControl7)
p_7 <- predict(fit_7, newdata = test7)
RMSE_test_7 <- RMSE(p_7, test7$y)
RMSE_test_7 # 4.361987
fit_7$bestTune # mtry = 2
set.seed(1, sample.kind = "Rounding")
fit_14 <- train(y ~ ., 
            data = data14,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid14,
            trControl=trControl14)
p_14 <- predict(fit_14, newdata = test14)
RMSE_test_14 <- RMSE(p_14, test14$y)
RMSE_test_14 # 4.620515
fit_14$bestTune # mtry = 13
set.seed(1, sample.kind = "Rounding")
fit_28 <- train(y ~ ., 
            data = data28,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid28,
            trControl=trControl28)
p_28 <- predict(fit_28, newdata = test28)
RMSE_test_28 <- RMSE(p_28, test28$y)
RMSE_test_28 # 4.716698
fit_28$bestTune # mtry = 11
set.seed(1, sample.kind = "Rounding")
fit_56 <- train(y ~ ., 
            data = data56,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid56,
            trControl=trControl56)
p_56 <- predict(fit_56, newdata = test56)
RMSE_test_56 <- RMSE(p_56, test56$y)
RMSE_test_56 # 5.939163
fit_56$bestTune # mtry = 8
set.seed(1, sample.kind = "Rounding")
fit_100 <- train(y ~ ., 
            data = data100,
            preProcess = c("center","scale"),
            method='parRF', 
            tuneGrid=tuneGrid100,
            trControl=trControl100)
p_100 <- predict(fit_100, newdata = test100)
RMSE_test_100 <- RMSE(p_100, test100$y)
RMSE_test_100 # 5.851088
fit_100$bestTune # mtry = 8
```

# Results
The test *RMSE* for each model in ascending order of age was `r round(RMSE_test_3,2)`, `r round(RMSE_test_7,2)`, `r round(RMSE_test_14,2)`, `r round(RMSE_test_28,2)`, `r round(RMSE_test_56,2)` and `r round(RMSE_test_100,2)` respectively. The table \ref{tab:table-reg-models} presents the details and results of each model (\ref{show-table-reg-models}). The figure \ref{fig:results-comparison} compares the actual and predicted values (\ref{show-results-comparison}), and the following tables show the best and worst results for each model (\ref{show-table-10}).
```{r table-reg-models, echo=F}
# Table - Models details
colNames <- c("Model", "mtry", "CV", "Repetitions", 
              "RMSE (train)", "RMSE (test)")
caption <- "Regression models results"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
dat_reg_models <- data.frame(
  dia = day,
  mtry = c(fit_3$bestTune$mtry, fit_7$bestTune$mtry, fit_14$bestTune$mtry,
           fit_28$bestTune$mtry, fit_56$bestTune$mtry, fit_100$bestTune$mtry),
  number = c(trControl3$number, trControl7$number, trControl14$number,
             trControl28$number,trControl56$number,trControl100$number),
  repeats = c(trControl3$repeats, trControl7$repeats, trControl14$repeats,
             trControl28$repeats,trControl56$repeats,trControl100$repeats),
  RMSE_train = c(min(fit_3$results$RMSE), min(fit_7$results$RMSE),
                  min(fit_14$results$RMSE), min(fit_28$results$RMSE),
                  min(fit_56$results$RMSE), min(fit_100$results$RMSE)),
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```
```{r results-comparison, echo=F, message=F, warning=F, fig.height=4, fig.cap=cap}
# Figure - Models comparison
cap <- "Actual vs Predicted values in each model"
models <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
xlabel <- "Actual (MPa)"
ylabel <- "Predicted (MPa)"
preds <- list(p_3, p_7, p_14, p_28, p_56, p_100)
actuals <- list(test3$y, test7$y, test14$y, test28$y, test56$y, test100$y)
gen_res_df <- function(ind){
  data.frame(actual = actuals[[ind]], pred = preds[[ind]], model = models[[ind]])
}
res_df <- lapply(1:6, gen_res_df)
res_df <- bind_rows(res_df)
res_df$model <- factor(res_df$modelo,levels=modelos)
res_df %>%
  ggplot(aes(actual, pred)) +
  facet_wrap(~ modelo, ncol=3) +
  geom_point(alpha=0.5) +
  theme_bw() +
  geom_abline(slope=1, intercept=0) +
  xlab(xlabel) +
  ylab(ylabel)
```
```{r table-10, echo=F}
# Results tables of 10 best and worst results
colNames1 = c("Actual", "Predicted", "Error", "","Actual", "Predicted", "Error") 
colNames2 = c("Best 10"=3,"", "Worst 10"=3) 
caption_0 <- "Model of "
get_X <- function(pred, actual, X=10){
  diff <- abs(pred - actual)
  diff_2 <- pred - actual
  ind_min <- which(diff <= max(sort(diff, decreasing = F)[1:X]), arr.ind = T)
  ind_max <- which(diff >= min(sort(diff, decreasing = T)[1:X]), arr.ind = T)
  df_min <- data.frame(
    actual_min=actual[ind_min],
    pred_min=pred[ind_min],
    diff_min=diff[ind_min],
    diff_min_2=diff_2[ind_min]
  )
  df_max <- data.frame(
    actual_max=actual[ind_max],
    pred_max=pred[ind_max],
    diff_max=diff[ind_max],
    diff_max_2=diff_2[ind_max]
  )
  df_max <- df_max[order(-df_max$diff_max),]
  df_min <- df_min[order(df_min$diff_min),]
  df_null <- data.frame(null_col = rep(c(""), X))
  res <- cbind(df_min,df_null, df_max)
  res <- res %>% select(-c(diff_max, diff_min))
  rownames(res) <- c()
  res
}
gen_kable <- function(ind){
  df <- get_X(preds[[ind]], actuals[[ind]])
  caption <- paste0(caption_0,models[ind])
  kable(
    df,
    col.names = colNames1,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  column_spec(4, width = "1cm",) %>%
  add_header_above(header = colNames2, line = T, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position"))
}
gen_kable(1)
gen_kable(2)
gen_kable(3)
gen_kable(4)
gen_kable(5)
gen_kable(6)
```

# Discussion
The models built present satisfactory results and prove that the compressive strength of concrete can be predicted relatively easily. The alternative adopted to create a model for each set of age proved to be a valid method, managing to stratify to obtain specific results for each set. The studies cited in the introduction using the same dataset have similar results, as expected. The \ref{tab:works-comparison} table presents the results of these works (\ref{show-works-comparison}), and the table \ref{tab:results} presents the values found for easy comparison (\ref{show-results}) .
```{r works-comparison, echo=F}
# Table - Comparison of other works
colNames = c("Author","Year" ,"Algorithm", "RMSE") 
caption <- "Comparison of other works"
works <- data.frame(
  autor = c("Pierobon", "Hameed", "Raj","Modukuru" ,"Alshamiri", "Abban"),
  ano = c("2018", "2020", "2018","2020" ,"2020", "2016"),
  modelo = c("Ensemble com 5 algorítimos", "Artificial Neural Networks", 
          "Gradient Boosting Regressor","Random Forest Regressor" ,
          "Regularized Extreme Learning Machine",
          "Support Vector Machines with Radial Basis Function Kernel"),
  RMSE = c("4.150", "4.736", "4.957","5.080","5.508", "6.105")
)
kable(
  works,
  col.names = colNames,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "l"
  )  %>%
    kable_styling(latex_options = c("HOLD_position"))
```
```{r results, echo=F}
# Table - Final results
colNames <- c("Model", "RMSE")
caption <- "Final result"
day <- c("3 days", "7 days", "14 days", "28 days", "56 days", "100 days")
dat_reg_models <- data.frame(
  dia = day,
  RMSE_test = c(RMSE_test_3, RMSE_test_7, RMSE_test_14,
                RMSE_test_28, RMSE_test_56, RMSE_test_100)
)
kable(
    dat_reg_models,
    col.names = colNames,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    )  %>%
  kable_styling(latex_options = c("HOLD_position"))
```
Following the line of reasoning of this work, it can be performed with different algorithms, the results found here used only one (*Parallel Random Forest*), even though it was theoretically the "best" found, other algorithms can present even better results. Another option is to create an *ensemble* of various algorithms, just like @Pierobon2018, but with the separation of age sets proposed here. In addition, it can be performed with a larger dataset, ideally with the same number of samples in each age set, a more homogeneous distribution of compressive strength, and less variance between samples.

# Literature cited
<div id="refs"></div>
\newpage

# Appendix 1 - Virtual environment {#appendix1}

## Operational system
```{r sistema, echo=FALSE}
v <- as.matrix(R.Version())
kable(v, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```

## Packages
```{r pacotes, echo=FALSE}
p <- matrix(
      c(
        "caret", as.character(packageVersion("caret")),
        "cowplot",as.character(packageVersion("cowplot")),
        "dplyr", as.character(packageVersion("dplyr")),
        "factoextra", as.character(packageVersion("factoextra")),
        "gdata", as.character(packageVersion("gdata")),
        "ggplot2", as.character(packageVersion("ggplot2")),
        "gridExtra",as.character(packageVersion("gridExtra")),
        "kableExtra",as.character(packageVersion("kableExtra")),
        "knitr",as.character(packageVersion("knitr")),
        "pastecs",as.character(packageVersion("pastecs")),
        "purrr", as.character(packageVersion("purrr")),
        "questionr", as.character(packageVersion("questionr")),
        "reshape2", as.character(packageVersion("reshape2")),
        "tidyr", as.character(packageVersion("tidyr")),
        "tidyverse", as.character(packageVersion("tidyverse"))
      ),
      ncol = 2,
      byrow=TRUE
    )
kable(p, booktabs = T) %>%
    kable_styling(latex_options = c("HOLD_position"))
```
\newpage

# Appendix 2 - Online repository {#appendix2}
All the code used to generate this work is available on *GitHub* at the following link:
- https://github.com/pedrobern/concrete-compressive-strength-prediction
\newpage

# Appendix 3 - Code {#appendix3}

## Obtaining the data
### Data download

```{r show-download-data, ref.label='download-data', ref='show-download-data', eval=FALSE}
```

### Ranaming the columns

```{r show-rename-dat-cols, ref.label='rename-dat-cols', ref='show-rename-dat-cols', eval=FALSE}
```

### Reordering the data

```{r show-reorder-dat, ref.label='reorder-dat', ref='show-reorder-dat', eval=FALSE}
```

### Defining column names and units

```{r show-col-names-and-units, ref.label='col-names-and-units', ref='show-col-names-and-units', eval=FALSE}
```

### Table First samples

```{r show-first-samples, ref.label='first-samples', ref='show-first-samples', eval=FALSE}
```

## Data preparation
### Removing duplicated samples

```{r show-removing-duplicated-samples, ref.label='removing-duplicated-samples', ref='show-removing-duplicated-samples', eval=FALSE}
```

### Tabela - Samples with same composition

```{r show-similar-samples, ref.label='similar-samples', ref='show-similar-samples', eval=FALSE}
```

### Tabela - Same samples with different results

```{r show-similar-samples-2, ref.label='similar-samples-2', ref='show-similar-samples-2', eval=FALSE}
```

### Data cleaning

```{r show-initial-data-cleaning, ref.label='initial-data-cleaning', ref='show-initial-data-cleaning', eval=FALSE}
```

### Tabela - Previous samples after processing

```{r show-similar-samples-same-id, ref.label='similar-samples-same-id', ref='show-similar-samples-same-id', eval=FALSE}
```

### Figure - Compressive strength (MPa) vs age (days)

```{r show-boxplot, ref.label='boxplot', ref='show-boxplot', eval=FALSE}
```

### Figure - Principal component analysis - 90, 91 e 100 days

```{r show-pca-90-91-100, ref.label='pca-90-91-100', ref='show-pca-90-91-100', eval=FALSE}
```

### Figure - Compressive strength through time

```{r show-mpa-on-time, ref.label='mpa-on-time', ref='show-mpa-on-time', eval=FALSE}
```

### Joining 90, 91 and 100 days data

```{r show-join-90-91-100, ref.label='join-90-91-100', ref='show-join-90-91-100', eval=FALSE}
```

### Figure - Ages frequency

```{r show-freq-ages, ref.label='freq-ages', ref='show-freq-ages', eval=FALSE}
```

### Removing ages with frequency lower than 50

```{r show-remove-ages-lower-50, ref.label='remove-ages-lower-50', ref='show-remove-ages-lower-50', eval=FALSE}
```

### Data reorganization

```{r show-reorganizing-dat, ref.label='reorganizing-dat', ref='show-reorganizing-dat', eval=FALSE}
```

## Data visualization
## Machine learning models
### Dummy variables

```{r show-dummy-var, ref.label='dummy-var', ref='show-dummy-var', eval=FALSE}
```

### Data preparation

```{r show-preparation, ref.label='preparation', ref='show-preparation', eval=FALSE}
```

### Table - First 18 columns of the 6 first samples of 28 days

```{r show-table-preparated-samples, ref.label='table-preparated-samples', ref='show-table-preparated-samples', eval=FALSE}
```

### Removing near zero variance columns

```{r show-nzv, ref.label='nzv', ref='show-nzv', eval=FALSE}
```

### High correlation variables verification

```{r show-cors, ref.label='cors', ref='show-cors', eval=FALSE}
```

### Test and train data split

```{r show-split, ref.label='split', ref='show-split', eval=FALSE}
```

### Distribution of test and train data

```{r show-dist-split, ref.label='dist-split', ref='show-dist-split', eval=FALSE}
```

### Naive model

```{r show-naive-model-reg, ref.label='naive-model-reg', ref='show-naive-model-reg', eval=FALSE}
```

### Table - Naive models

```{r show-table-naive-model-reg, ref.label='table-naive-model-reg', ref='show-table-naive-model-reg', eval=FALSE}
```

### Features selection

```{r show-test-models-reg, ref.label='test-models-reg', ref='show-test-models-reg', eval=FALSE}
```

### Table - First 6 samples of train data of the 28 days model

```{r show-table-done-samples-28, ref.label='table-done-samples-28', ref='show-table-done-samples-28', eval=FALSE}
```

### Regression models

```{r show-reg-models, ref.label='reg-models', ref='show-reg-models', eval=FALSE}
```

# Results
### Table - Models details

```{r show-table-reg-models, ref.label='table-reg-models', ref='show-table-reg-models', eval=FALSE}
```

### Figure - Models comparison

```{r show-results-comparison, ref.label='results-comparison', ref='show-results-comparison', eval=FALSE}
```

### Results tables of 10 best and worst results

```{r show-table-10, ref.label='table-10', ref='show-table-10', eval=FALSE}
```

# Discussion
### Table - Comparison of other works

```{r show-works-comparison, ref.label='works-comparison', ref='show-works-comparison', eval=FALSE}
```

### Table - Final results

```{r show-results, ref.label='results', ref='show-results', eval=FALSE}
```